Stellar Mass and Stellar Kinematics Can Constrain Gravitational Lens Models

### 2014-07-23 limits/ms.tex
Changes mainly to the refreee reply from Justin; not yet finished.

@@

To see why lens modelling details are of crucial importance, let us
recall the essential quantities that appear in lensing (see also \S\ref{sec:theory} for a more detailed exposition). First we have the distances. Let $D_L$, $D_S$, $D_{LS}$ be the
angular-diameter distances to the lens, source, and from lens to
source; these are all proportional to $c/H_0$ but have factors
that depend on the particular choice of cosmology\footnote{Here, $c$ is the speed of light in vacuo and $H_0$ is the Hubble parameter.}.  Typically:
%
\begin{equation}
D_L \approx z_L \frac c{H_0}\ \hbox{and}\ \frac{D_S}{D_{LS}} \sim 1.
\end{equation}
where $z_L$ is the redshift of the lens. 
%
For multiple images, the sky-projected density must exceed the critical
density in some region:
%
\begin{equation}
\Sigma_\mathrm{crit} = \frac{c^2}{4\pi G D_L} \sim \frac{1\rm\,kg\,m^{-2}}{z_L}
\end{equation}
where $G$ is Newton's gravitational constant. 
%
The angular separation between the lensed images is of order
the Einstein radius $\theta_E$, which is related to
the mass by:
%
\begin{equation}
\theta_E \sim \frac{R_G}{D_L} \frac{D_{LS}}{D_S}
\end{equation}
%
where $R_G = GM/c^2$ (with $M$ the projected mass enclosed within $\theta_E$)
is the gravitational radius. If the source is a quasar or otherwise rapidly
variable, a time delay $\Delta t$ in the variability will be present where
%
\begin{equation}
\Delta t \sim R_G/c
\end{equation}
%
So in principle, one can not only measure the mass of the lens, one can use the
dependence on the cosmology-dependent $D$ factors to extract the cosmological
model and all its parameters.  \cite{1937ApJ....86..217Z} drew attention to the
former, and \cite{1964MNRAS.128..307R,1966MNRAS.132..101R} pointed out the
latter, all long before lenses were discovered.  The difficulty with actually
doing this, however, became apparent soon after the discovery of the first lens
by \cite{1979Natur.279..381W}.  In the first ever paper on lens modelling,
\cite{1981ApJ...244..736Y} found that many plausible mass distributions could
reproduce the data.

>>>>

\keywords{gravitational lensing: strong, methods: numerical, methods: statistical}

@@

To see why lens modelling details are of crucial importance, let us
recall the essential quantities that appear in lensing (see also \S\ref{sec:theory} for a more detailed exposition). First we have the distances. Let $D_L$, $D_S$, $D_{LS}$ be the
angular-diameter distances to the lens, source, and from lens to
source; these are all proportional to $c/H_0$ but have factors
that depend on the particular choice of cosmology\footnote{Here, $c$ is the speed of light in vacuo and $H_0$ is the Hubble parameter.}.  Typically:
%
\begin{equation}
D_L \approx z_L \frac c{H_0}\ \hbox{and}\ \frac{D_S}{D_{LS}} \sim 1.
\end{equation}
where $z_L$ is the redshift of the lens. 
%
For multiple images, the sky-projected density must exceed the critical
density in some region:
%
\begin{equation}
\Sigma_\mathrm{crit} = \frac{c^2}{4\pi G D_L} \sim \frac{1\rm\,kg\,m^{-2}}{z_L}
\end{equation}
where $G$ is Newton's gravitational constant. 
%
The angular separation between the lensed images is of order
the Einstein radius $\theta_E$, which is related to
the mass by:
%
\begin{equation}
\theta_E \sim \frac{R_G}{D_L} \frac{D_{LS}}{D_S}
\end{equation}
%
where $R_G = GM/c^2$ (with $M$ the projected mass enclosed within $\theta_E$)
is the gravitational radius. If the source is a quasar or otherwise rapidly
variable, a time delay $\Delta t$ in the variability will be present where:
%
\begin{equation}
\Delta t \sim R_G/c
\end{equation}
%
So in principle, one can not only measure the mass of the lens, one can use the
dependence on the cosmology-dependent $D$ factors to extract the cosmological
model and all its parameters.  \cite{1937ApJ....86..217Z} drew attention to the
former, and \cite{1964MNRAS.128..307R,1966MNRAS.132..101R} pointed out the
latter, all long before lenses were discovered.  The difficulty with actually
doing this, however, became apparent soon after the discovery of the first lens
by \cite{1979Natur.279..381W}.  In the first ever paper on lens modelling,
\cite{1981ApJ...244..736Y} found that many plausible mass distributions could
reproduce the data.
